Stack Safety is a Security Property

### 2021-04-17 tex/main.tex
Answering notes and simplifying

Finally, we formalize a version of WBCF, which captures the same informal idea
as that of \citet{SkorstengaardSTK} using our context machinery.
\definition
A system enjoys \textit{\textbf{well-bracketed control flow}} when, for every pair of reachable states
\(\mpcstatename = (\underscore,\underscore,(\underscore,\rts))\) and
\(\mpcstatename' = (\underscore,\underscore,(\underscore,\rts'))\) such that
\(\mpcstatename \stepsto \mpcstatename'\)\apt{doesn't reachability of $\mpcstatename$ imply reachability of $\mpcstatename'$?}, if \(|\rts'| < |\rts|\), then
\(|\rts'| = |\rts|-1\).

@@

More interesting changes to the tags (from a stack safety perspective) 
happen at exactly the points in the program where we switch from a
caller to a callee or vice-versa:

@@

\apt{Given that the deferred property suffices to specifcy the lazy policies,
  the case for observable properties is pretty weak.  Should we just cut them?}

@@

\paragraph{Formalization of variable passing.}
We add extra information to our call
annotations to reflect that they may now pass variables on the stack, and that the locations
of those variables may be dynamic (as in the case of pass-by-reference).
A call annotation now includes a predicate relating machine states to the components
that should be passed---when passing by value, this will typically be freshly allocated
cells at the top of the frame, while passing by reference just gives the
addresses of the appropriate variables relative to the stack pointer.
\apt{This still doesn't make sense to me. We cannot
  give a static annotation that identifies the absolute address of a stack-allocated variable!
  Only sp-relative addresses make sense.}\leo{Better now? :)}

\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(P) \hspace*{1in} P \subseteq (\MACHS \times \COMPONENTS)\]
We extend the set of domains to include passed components,
which are labeled with the range of depths of the caller that has passed them.
\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \passed{\depth,\depth'}\]
%\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
%\times (\rts \in \listT (\target \subseteq \MACHS)) \]
The definition of context looks the same, but is based on the revised definition of \(\STKDOMS\):
\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
 \times (\listT (\target)) \]
We straightforwardly describe how the sealing convention and passing predicate combine to
update a domain map at a call from depth \(\depth\).
\[(\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) ~ \component =
\begin{cases}
  \sealed{\depth}              & \text{if } \domMap ~ \component = \unsealed \text{ and }
                                 \sealcon ~ \mach ~ \component \text{ and } \neg P ~ \mach ~ \component \\
  \passed{\depth,\depth+1}     & \text{if } \domMap ~ \component = \unsealed \text{ and }
                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component \\
  \passed{\depth',\depth+1}    & \text{if } \domMap ~ \component = \passed{\depth',\depth} \text{ and }
                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component,
                                 \text{ for any } \depth' \\
  \domMap~\component & \text{else} \\
\end{cases}\]

@@

\definition
For each of the stepwise, deferred, and observational versions of
stack integrity in the simple model, the equivalent property in the
passing model substitutes the above definitions of \(\context\), initial contexts\apt{not defined above??}  and update
function, and \(\prot_P\) for their simple equivalents, and is otherwise identical.

@@

\definition
For both the stepwise and observational versions of stack confidentiality in
the original model, the equivalent property in the passing model is the same after
substituting the above definitions of \(\context\), initial contexts\apt{ditto} and update
function, and \(\prot_P\) for their simple equivalents.

>>>>

Finally, we formalize a version of WBCF, which captures the same informal idea
as that of \citet{SkorstengaardSTK} using our context machinery.
\definition
A system enjoys \textit{\textbf{well-bracketed control flow}} when, for every reachable state
\(\mpcstatename\) where \(\pi_\context(\mpcstatename) = (\underscore,\rts))\), and
\(\mpcstatename \stepsto \mpcstatename'\) where \(\pi_\context(\mpcstatename') = (\underscore,\rts'))\),
if \(|\rts'| < |\rts|\), then \(|\rts'| = |\rts|-1\).

@@

More interesting (from a stack safety perspective) changes to the tags
happen at exactly the points in the program where we switch from a
caller to a callee or vice-versa:

@@

\ifaftersubmission
\apt{Given that the deferred property suffices to specifcy the lazy policies,
  the case for observable properties is pretty weak.  Should we just cut them?}
\fi

@@

\paragraph{Formalization of variable passing.}
We add extra information to our call
annotations to reflect that they may now pass variables on the stack, and that the locations
of those variables may be dynamic (as in the case of pass-by-reference).
A call annotation now includes a predicate relating machine states to the components
that should be passed---we can always determine where they should be in terms of their
offsets from the stack pointer, and embed this knowledge in an annotation.

\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(P) \hspace*{1in} P \subseteq (\MACHS \times \COMPONENTS)\]
We extend the set of domains to include passed components,
which are labeled with the range of depths of the caller that has passed them.
\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \passed{\depth,\depth'}\]
The definition of context looks the same, but is based on the revised definition of \(\STKDOMS\):
\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
\times (\listT (\target)) \]
The initial context again maps all components to \(\unsealed\).
We straightforwardly describe how the sealing convention and passing predicate combine to
update a domain map at a call from depth \(\depth\).
\[(\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) ~ \component =
\begin{cases}
  \sealed{\depth}              & \text{if } \domMap ~ \component = \unsealed \text{ and }
                                 \sealcon ~ \mach ~ \component \text{ and } \neg P ~ \mach ~ \component \\
  \passed{\depth,\depth+1}     & \text{if } \domMap ~ \component = \unsealed \text{ and }
                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component \\
  \passed{\depth',\depth+1}    & \text{if } \domMap ~ \component = \passed{\depth',\depth} \text{ and }
                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component,
                                 \text{ for any } \depth' \\
  \domMap~\component & \text{else} \\
\end{cases}\]

@@

@@

\definition
For each of the stepwise, deferred, and observational versions of
stack integrity in the simple model, the equivalent property in the
passing model substitutes the above \(\CONTEXTS\) type and update
function, and \(\prot_P\), for their equivalents in the simple model.

For both the stepwise and observational versions of stack confidentiality in
the original model, the passing model equivalents additionally substitute \(\secret_P\).
